---
title: "song lyric analysis"
author: "kp"
date: "29 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
library(tidytext) # for tf-idf matrix
library(topicmodels) # for LDA
library(ggplot2)
```

## Import data
Read csv of song lyrics and view subset:
```{r}
dat <<- read.csv("./songdata.csv", strings = FALSE)
#head(dat)
#tail(dat)
summary(dat$text)
```

## Data Clean/Exploration
Data looks fairly clean aside from some misspelling that I'm not going to bother dealing with.

Look at number of artists that appear frequently.

```{r}
dat <- dat %>% 
  mutate(sa = paste0(artist, "_", song)) %>% 
  group_by(artist) %>% 
  mutate(n = n())

table(dat$artist[dat$n > 180])
```



Remove one-hit wonders
```{r}
dat <- dat[dat$n > 10,]
```

split into words
```{r} 
by_word<- dat %>%
  unnest_tokens(word, text) 

by_word$link <- NULL
by_word$n <- NULL

word_counts <- by_word %>%
  anti_join(stop_words) %>%
  count(sa, word, sort = TRUE) %>%
  ungroup()

```

## LDA on songs

Get document-term matrix
```{r}
songs_dtm <- word_counts %>%
  cast_dtm(sa, word, n)

songs_dtm
```

Run LDA

```{r}
# try 8 groupings to start, though suspect this is too high - each run takes ~30 mins on full data
songs_lda <- LDA(songs_dtm, k = 8, control = list(seed = 1234))

```

look at probabilities
```{r}
song_topics <- tidy(songs_lda, matrix = "beta")

# take a guess at possible topics
song_topics[song_topics$term %in% "love",] # topic 4

```


Top terms per topic
```{r}
top_terms <- song_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

top_terms
```
  
```{r}
# plot top 8 terms per topic

top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```

# look at documents per group (artists/songs)

```{r}
songs_gamma <- tidy(songs_lda, matrix = "gamma")

# for each song, assign topic with highest gamma value
topic_song <- songs_gamma %>% 
  group_by(document) %>% 
  arrange(desc(gamma)) %>% 
  slice(1) # take top topic

by_artist <- topic_song %>% 
  separate(col = document, into = c("artist", "song"), sep = "_")
# look at lyric diversity by artist
hist(by_artist$topic[by_artist$artist %in% "Adele"])
hist(by_artist$topic[by_artist$artist %in% "Outkast"])
hist(by_artist$topic[by_artist$artist %in% "Talking Heads"])
hist(by_artist$topic[by_artist$artist %in% "Queen"])
hist(by_artist$topic[by_artist$artist %in% "Cher"])
hist(by_artist$topic[by_artist$artist %in% "Adam Sandler"]) # ???

```

  
# next steps - do one just beatles - webscrape more varied songs - predictive 